Computable Poincar\'e--Friedrichs constants for the $L^{p}$ de Rham complex over convex domains and domains with shellable triangulations

### 2024-10-05 main.tex
singular values inserted

We return to the theory of triangulations,
as our main objective requires some further concepts.
We are interested in simplicial complexes that triangulate manifolds. 
Our particular interest are manifold triangulations that are \emph{shellable}.
Such simplicial complexes are constructed by successively adding simplices in a particularly well-behaved manner. 
Local patches (stars) within triangulations of dimension two or three are examples of such shellable complexes. 
The monographs by Kozlov~\cite{kozlov2008combinatorial} and Ziegler~\cite{ziegler1995lectures} are our main references for this section. 
We also refer to Lee's monograph~\cite{lee2011topological} for any further background on manifolds.

@@

The following special cases receive particular interest:
an \textit{$n$-ball triangulation} is any triangulation of a topological (closed) $n$-ball\mwl{, and we sometimes call this a \textit{$n$-disk} triangulation}.
An \textit{$n$-sphere triangulation} is any triangulation of a topological $n$-sphere.

@@

We say that an $n$-dimensional manifold triangulation $\calT$ is \emph{combinatorial}
if for every vertex $V \in \calT$ the local star $\patch_{\calT}(V)$ is a simplicial ball of dimension $n$.

\begin{lemma}\label{lemma:startopology}
    Let $\calT$ be a finite $n$-dimensional simplicial complex whose underlying space is a manifold $\calM$.
    Suppose that $1 \leq n \leq 3$. Then the following holds:
    \begin{itemize}
        \item
        If $S \in \calT$ is an inner simplex, 
        then $\patch_{\calT}(S)$ is a simplicial $n$-ball with $S$ as an inner simplex
        and $\carapace_{\calT}(S)$ is a simplicial $(n-1)$-sphere. 
        \item
        If $S \in \calT$ is a boundary subsimplex, 
        then $\patch_{\calT}(S)$ is a simplicial $n$-ball with $S$ as a boundary simplex,
        and $\carapace_{\calT}(S)$ is a simplicial $(n-1)$-ball.
    \end{itemize}
\end{lemma}
\begin{proof}
    The lemma is obvious if $n = 1$, so we assume $n \geq 2$ in what follows. 
    We prove these statements in several steps. 
    The reader is assumed to have some background in topology. 
    \begin{enumerate}
    \item 
    Let $S$ be any simplex with vertices $v_0, v_1, \dots, v_k$, with barycenter $z_{S}$, and dimension $k$.
    Let $\calS := \patch_{\calT}(S)$ be its star. 
    Each $l$-dimensional simplex $T \in \calS$ that contains $S$ 
    has vertices $v_0, v_1, \dots, v_k, v_{k+1}^{S}, \dots, v_{l}^{S}$. 
    For any such simplex, we introduce a decomposition $T_{0}, \dots, T_{k}$, where each $T_{i}$ has vertices 
          % $v_0, \dots, v_{i-1}, z_{S}, v_{i+1}, \dots, v_k, v_{k+1}^{S}, \dots, v_{l}^{S}$.
    \begin{align*}
        v_0, \dots, v_{i-1}, z_{S}, v_{i+1}, \dots, v_k, v_{k+1}^{S}, \dots, v_{l}^{S}. 
    \end{align*}
    The collection $\calS'$ of these simplices and their subsimplices constitute a simplicial complex 
    that triangulates the same underlying set as $\calS$.
    Moreover, $\calS' = \patch_{\calS'}(z_{S})$. 
    In particular, $z_{S}$ is a boundary vertex of $\calS'$ if and only if $S$ is a boundary simplex of $\calS$. 
    So it remains to study the topology of vertex stars.

@@

\begin{example}\label{example:shell_simpl}
    Any simplex $T$ (trivially) has a shelling. The boundary complex $\partial \calT(T)$ has a shelling:
    in fact, any enumeration of the boundary faces of $T$ constitutes a shelling\mwl{; see Example 8.2.(iii) in~\cite{ziegler1995lectures}}.
\end{example}
\begin{example}
    The standard triangulation of the $3$-dimensional cube by six tetrahedra, the Kuhn triangulation~\cite{kuhn1960some}, is shellable, as are its higher-dimensional generalizations.\footnote{We remark that Kuhn attributes this triangulation to Lefschetz~\cite{lefschetz2015introduction}.}
\end{example}
\begin{example}
    There exists a non-shellable triangulation of a tetrahedron and of a cube in $n=3$, see~\cite[Example 8.9]{ziegler1995lectures}.
\end{example}

@@

    Clearly, this is the case if the triangulation of the $2$-\disk\ contains two triangles. 
    Now suppose the induction claim is true when the triangulation includes at most $N$ triangles,
    and assume that $\calT$ includes $N+1$ triangles. 
    There are at least two triangles with an edge on the boundary. 
    Suppose that $\calT$ does not have at least two triangles that are good in $\calT$.
    Then there exists a triangle $T'$ that intersects $\partial M$ 
    in one edge and its opposite vertex.
    Removing $T'$ splits the manifold into two face-connected components, each of which is a topological $2$-\disk.
    By the induction assumption, each of those components contains at least two triangles 
    that are good in the respective component. 
    So each component has at least one triangle that is good in $\calT$. 
    Hence $\calT$ contains two good triangles, which completes the induction step.

@@

\begin{remark}
    Not all triangulable sets admit a triangulation that is shellable. 
    Moreover, even if a set admits a shellable triangulation, not all of its triangulations are be shellable. For example, if we extend the non-shellable triangulation of the tetrahedron from~\cite[Example 8.9]{ziegler1995lectures} to a triangulation of a hypertetrahedron by suspending it from a new point $v_\star$, then the resulting new triangulation is non-shellable and coincides with the patch around $v_\star$.
    This demonstrates that patches around boundary simplices are not necessarily shellable when the space dimension $n$ is larger than three.
\end{remark}
\begin{remark}
    Not all triangulable sets admit a triangulation that is also shellable. 
    Moreover, even if a set admits a shellable triangulation, not all of its triangulations might be shellable. 
    We refer to~\cite[Example 8.9]{ziegler1995lectures} for an examples of non-shellable triangulations of cubes and tetrahedra in three dimensions. 
\end{remark}

@@

\begin{proposition}\label{proposition:starreflection}
    Let $\calT$ be an $n$-dimensional triangulation of an $n$-dimensional domain. 
    Let $S \in \calT$ be a simplex of dimension $k < n$,
    let $T \in \patch_{\calT}(S)$ be of dimension $n$,
    and let 
    \begin{align*}
        U := \overline{ |\patch_{\calT}(S)| \setminus T },
        \qquad 
        {\Gamma_1} := U \cap T,
        \qquad 
        {\Gamma_2} := \overline{ \partial T \setminus \partial U }.
    \end{align*}
    The following holds, where the constants on the right-hand sides are as stated in the proof. 
    \begin{enumerate}
    \item 
    There exists a bi-Lipschitz mapping
    \begin{align*}
        \Xi_{1} : T \rightarrow U
    \end{align*}
    which is the identity along ${\Gamma_1} = T \cap U$. 
    Its singular values $\sigma_1 \geq \dots \geq \sigma_n$ satisfy 
    \begin{align*}
        \sigma_1 &\leq ...
        \\
        \sigma_i &\leq ..., \quad 2 \leq i \leq n-1,
        \\
        \sigma_n &\leq ...
    \end{align*}
    and which satisfies the upper bounds 
    \begin{gather*}
        \matnorm{ \Jacobian\Xi_{1} }          \leq \Cfive{n}{k}(\calT),
        \quad 
        \matnorm{ \Jacobian\Xi_{1}^{-1} }     \leq \Csechs{n}{k}(\calT),
        \\
        | \det\Jacobian\Xi_{1} |      \leq \Cdetfive{n}{k}(\calT),
        \quad 
        | \det\Jacobian\Xi_{1}^{-1} | \leq \Cdetsechs{n}{k}(\calT).
    \end{gather*}
    \item 
    There exists a bi-Lipschitz mapping
    \begin{align*}
        \Xi_{2} : \underlying{\patch_{\calT}(S)} \rightarrow U
    \end{align*}
    which is the identity along $\partial U \setminus \partial T$.
    Its singular values $\sigma_1 \geq \dots \geq \sigma_n$ satisfy 
    \begin{align*}
        \sigma_1 &\leq ...
        \\
        \sigma_i &\leq ..., \quad 2 \leq i \leq n-1,
        \\
        \sigma_n &\leq ...
    \end{align*}
    and which satisfies the upper bounds 
    \begin{gather*}
        \matnorm{ \Jacobian\Xi_{2} }          \leq \Csieb{n}{k}(\calT),
        \quad 
        \matnorm{ \Jacobian\Xi_{2}^{-1} }     \leq \Cacht{n}{k}(\calT),
        \\
        | \det\Jacobian\Xi_{2} |      \leq \Cdetsieb{n}{k}(\calT),
        \quad 
        | \det\Jacobian\Xi_{2}^{-1} | \leq \Cdetacht{n}{k}(\calT).
    \end{gather*}
    \end{enumerate}
\end{proposition}
\mwl{TODO: throughout this manuscript, all vector and operator norms should indicate the exponent.}
\begin{proof}
    We derive the estimate in several steps. 
    In what follows, we use the notation $\hat z$ for the normalization of any vector $z \in \bbR^{n}$.
    \begin{itemize}
        \item 
        Without loss of generality, the barycenter $z_{S}$ of $S$ is the origin. 
        % Let $I := U \cap T$, which is the union of $k$ different faces of $T$, with $1 \leq k \leq n$.
        %We let ${\Gamma_2} \subseteq \partial T$ be the union of the remaining $n+1-k$ faces of $T$.
        %So ${\Gamma_2$} is within the boundary of the star $|\patch_{\calT}(S)|$ and is disjoint from $S$.
        We fix the subsimplex $S' \subseteq T$ that is complementary to $S$.
        Note that ${\Gamma_2}$ is the union of exactly those faces of $T$ that contain $S'$,
        whereas ${\Gamma_1} = U \cap T$ is the union of exactly those faces of $T$ that contain $S$.

@@

        \mwl{Check that $\calS'$ is a triangulation...}

@@

            \sigma_{\max} = \sqrt{ 2\tan(\beta)^2 + 1 + 2 \tan(\beta) \sqrt{ \tan(\beta)^2 + 1 } } = \sqrt{ 1 + \tan^{2}(\beta) } + \tan(\beta)

@@

            \sigma_{\min} = \sqrt{ 2\tan(\beta)^2 + 1 - 2 \tan(\beta) \sqrt{ \tan(\beta)^2 + 1 } } = \sqrt{ 1 + \tan^{2}(\beta) } - \tan(\beta)

@@

        Recall that $K \in \calR$ is obtained from $T$ via barycentric subdivisions, first at $z_{S}$ and then at $z_{S'}$. 
        Let $F_{z} \subseteq K$ be the face opposite to the vertex $z_{S'}$,
        which must be contained in some face of $T$. 
        Since $S'$ has dimension $n-k-1$, the vector $(n-k) h_{z}$ is the height vector of that face of $T$. 
        Thus 
        \begin{align*}
            \frac{ \vecnorm{ b_{z} } }{ \vecnorm{ h_{z} } }
            = 
            \frac{ \vecnorm{ b_{z} } }{ (n-k)^{-1} \| (n-k) h_{z} \| }
            = 
            \frac{ \vecnorm{ z_{S'} } }{ (n-k)^{-1} \| (n-k) h_{z} \| }
            \leq 
            (n-k) \aspectratio(T)
%             = 
%             \frac{ \vecnorm{ b_{z} } \vol(F_{z}) }{ (n-k)^{-1} n \vol(T) }
%             \leq 
%             \frac{ n-k }{n!}
%             \cdot 
%             \frac{ \diam(T)^{n} }{ \vol(T) }
            .
        \end{align*}
        This establishes bounds on the singular values of the transformation. 
        We abbreviate 
        \begin{align*}
            B_{T} 
            := 
%             \sqrt{ 1 + \tan^{2}(\beta) } - \tan(\beta) 
%             \geq 
            \sqrt{ 
                1 + (n-k)^{2} \aspectratio(T)^{2} % \frac{ \diam(T)^{2n} }{ n!^{2}\vol(T)^{2} } 
            } 
            + 
            (n-k) \aspectratio(T) % \frac{ \diam(T)^{n} }{ n! \vol(T) }            
            .
        \end{align*}

@@

@@

@@

@@

            \Psi : |\calR'| \rightarrow |\calK|.

@@

        \mwl{Here we need a particular result. }

@@

@@

@@

        The vector $a$ is the difference between $\hat h_{G}$ and $\hat h_{K}$ after both are rescaled such that they lie on the affine hyperplane with normal vector $\hat x$. That difference has length at most $\tan(\alpha_{G}) + \tan(\alpha_{K})$.
        We calculate 
        \begin{align*}
            \|a\|
            = 
            \left\|
            \frac{\|x_{G}\|}{\|h_{G}\|} \hat h_{{G}} 
            - 
            \frac{\|x_{K}\|}{\|h_{K}\|} \hat h_K
            \right\|
            &
            \leq 
            \frac{\|x_{G}\|}{\|h_{G}\|} 
            + 
            \frac{\|x_{K}\|}{\|h_{K}\|} 
            % \\&
            \leq 
            (k+1) 
            \left( 1 + B_{T} \right)
            \sup_{ K \in \patch_{\calT}(S) }
            \aspectratio(K) % \frac{ \diam(K)^{n} }{ n! \vol(K) }
            .
        \end{align*}

        We have shown that $\Psi$ and $\Psi^{-1}$ are piecewise Lipschitz with respect to some essentially disjoint closed partitions over their respective domains.
        It remains to show both are Lipschitz, which we illustrate for $\Psi$, the argument for $\Psi^{-1}$ being analogous. 
        Rademacher's theorem then shows that both have essentially bounded weak Jacobians.

        We show that every function that is piecewise Lipschitz with respect to the triangulation $\calK$ must be Lipschitz over all of $|\calK|$. 
        We use a technique similar to Lemma~\ref{lemma:stars_are_shellable}. 
        After application of a piecewise linear Lipschitz mapping, 
        we can assume that $|\calK|$ is convex without loss of generality. 
        Any function piecewise Lipschitz with respect to $\calK$ must be Lipschitz over $|\calK|$. 
        \mwl{TODO: not enough. Instead, use the homeomorphism onto a ball and estimate its Lipschitz constant.}

        \item 
        \mwl{Use interlacing inequalities.}
        We define a mapping $ \Xi_{1} : T \rightarrow U = \overline{\patch_{\calT}(S) \setminus T}$ by setting 
        \begin{align*}
            \Xi_{1} := \Psi \circ \Theta.
        \end{align*}
        It preserves the interface from $T$ onto $U$. 
        Let $\sigma_1(x) \geq \dots \geq \sigma_n(x)$ be the singular values of its Jacobian at $x \in T$. 
        % We briefly recapitulate why it is bi-Lipschitz.
%         We show that every function that is piecewise Lipschitz with respect to the triangulation $\calK$ must be Lipschitz over all of $|\calK|$. 
%         We use a technique similar to Lemma~\ref{lemma:stars_are_shellable}. 
%         After application of a piecewise linear Lipschitz mapping, 
%         we can assume that $|\calK|$ is convex without loss of generality. 
%         Any function piecewise Lipschitz with respect to $\calK$ must be Lipschitz over $|\calK|$. 
%         By an application of Rademacher's theorem, we conclude that $\calK$ is bi-Lipschitz. 
        The combination of the previous estimates for $\Psi$ and $\Theta$,
        as well as the upper bound $\sqrt{s^2+t^2} \leq s + t$ for any $s,t \geq 0$,
        provide 
        \begin{align*}
            \matnorm{ \Jacobian \Xi_{1} }
            &\leq 
            \Cfive{n}{k}(\calT)
            :=
            (k+1) 
            \diameterratio(\calT)
            \aspectratio(\calT)
            B_{T}^{2}
            \Big( 
                1 
                + 
                (k+1) 
                \left( 1 + B_{T} \right)
                \aspectratio(\calT)
            \Big)            
            ,
            \\
            \matnorm{ \Jacobian \Xi_{1}^{-1} }
            &\leq  
            \Csechs{n}{k}(\calT)
            :=
            (k+1) 
            \diameterratio(\calT)
            \aspectratio(\calT)
            B_{T}
            \Big( 
                1 
                + 
                (k+1) 
                \left( 1 + B_{T} \right)
                \aspectratio(\calT)
            \Big)            
            .
        \end{align*}
        We get the determinant estimates
        \begin{align*}
            |\det(\Xi_{1})|
            &\leq 
            \Cdetfive{n}{k}(\calT)
            :=
            (k+1)^{n}
            \diameterratio(\calT)^{n}
            B_{T}^{n}
            \aspectratio(\calT)^{n}
            ,
            \\
            |\det(\Xi_{1}^{-1})|
            &\leq 
            \Cdetsechs{n}{k}(\calT)
            :=
            (k+1)^{n}
            \diameterratio(\calT)^{n}
            \aspectratio(\calT)^{n}
            .
        \end{align*}
        This completes the desired estimates for $\Xi_{1}$.

>>>>

We return to the theory of triangulations, as our main objective requires some further concepts.
We are interested in simplicial complexes that triangulated domains and which are \emph{shellable}.
Such simplicial complexes are constructed by successively adding simplices in a well-structured manner. 
Local patches (stars) within triangulations of dimension two or three are examples of such shellable complexes. 
The monographs by Kozlov~\cite{kozlov2008combinatorial} and Ziegler~\cite{ziegler1995lectures} are our main references for this section. 
We also refer to Lee's monograph~\cite{lee2011topological} for any further background on manifolds.

@@

The following special cases receive particular interest:
an \textit{$n$-ball triangulation} is any triangulation of a topological (closed) $n$-ball, and we sometimes call this an \textit{$n$-disk} triangulation.
An \textit{$n$-sphere triangulation} is any triangulation of a topological $n$-sphere.

@@

\begin{lemma}\label{lemma:startopology}
    Let $\calT$ be a finite $n$-dimensional simplicial complex whose underlying space is a manifold $\calM$.
    Suppose that $1 \leq n \leq 3$. Then the following holds:
    \begin{itemize}
        \item
        If $S \in \calT$ is an inner simplex, 
        then $\patch_{\calT}(S)$ is a simplicial $n$-ball with $S$ as an inner simplex
        and $\carapace_{\calT}(S)$ is a simplicial $(n-1)$-sphere. 
        \item
        If $S \in \calT$ is a boundary subsimplex, 
        then $\patch_{\calT}(S)$ is a simplicial $n$-ball with $S$ as a boundary simplex,
        and $\carapace_{\calT}(S)$ is a simplicial $(n-1)$-ball.
    \end{itemize}
\end{lemma}
\begin{proof}
    The lemma is obvious if $n = 1$, so we assume $n \geq 2$ in what follows. 
    We prove these statements in several steps. 
    The reader is assumed to have some background in topology. 
    \begin{enumerate}
    \item 
    Let $S$ be any simplex with vertices $v_0, v_1, \dots, v_k$, with barycenter $z_{S}$, and dimension $k$.
    Let $\calS := \patch_{\calT}(S)$ be its star. 
    Each $l$-dimensional simplex $T \in \calS$ that contains $S$ 
    has vertices $v_0$, $v_1$, $\dots$, $v_k$, $v_{k+1}^{S}$, $\dots$, $v_{l}^{S}$. 
    For any such simplex, we introduce a decomposition $T_{0}, \dots, T_{k}$, where each $T_{i}$ has vertices 
          % $v_0, \dots, v_{i-1}, z_{S}, v_{i+1}, \dots, v_k, v_{k+1}^{S}, \dots, v_{l}^{S}$.
    \begin{align*}
        v_0, \dots, v_{i-1}, z_{S}, v_{i+1}, \dots, v_k, v_{k+1}^{S}, \dots, v_{l}^{S}. 
    \end{align*}
    The collection $\calS'$ of these simplices and their subsimplices constitute a simplicial complex 
    that triangulates the same underlying set as $\calS$.
    Moreover, $\calS' = \patch_{\calS'}(z_{S})$. 
    In particular, $z_{S}$ is a boundary vertex of $\calS'$ if and only if $S$ is a boundary simplex of $\calS$. 
    So it remains to study the topology of vertex stars.

@@

\begin{example}\label{example:shell_simpl}
    Any simplex $T$ (trivially) has a shelling, consisting only of $T$ itself. 
	The boundary complex $\partial \calT(T)$ has a shelling:
    any enumeration of the boundary faces of $T$ constitutes a shelling\mwl{; see Example 8.2.(iii) in~\cite{ziegler1995lectures}}.
\end{example}
\begin{example}
    The standard triangulation of the $3$-dimensional cube by six tetrahedra, the Kuhn triangulation~\cite{kuhn1960some}, is shellable, as are its higher-dimensional generalizations.\footnote{We remark that Kuhn attributes this triangulation to Lefschetz~\cite{lefschetz2015introduction}.}
\end{example}
\begin{example}
    There exists a non-shellable triangulation of a tetrahedron and of a cube in $n=3$, see~\cite[Example 8.9]{ziegler1995lectures}.
\end{example}

@@

    Clearly, this is the case if the triangulation of the $2$-\disk\ contains two triangles. 
    Now suppose the induction claim is true when the triangulation includes at most $N$ triangles,
    and assume that $\calT$ includes $N+1$ triangles. 
	As we travel along the boundary, we traverse along edges of at least two simplices, 
	and therefore there are at least two triangles with an edge on the boundary. 
    Suppose that $\calT$ does not have at least two triangles that are good in $\calT$.
    Then there exists a triangle $T'$ that intersects $\partial M$ 
    in one edge and its opposite vertex.
    Removing $T'$ splits the manifold into two face-connected components, each of which is a topological $2$-\disk. % TODO: why are those two disks?
    By the induction assumption, each of those components contains at least two triangles 
    that are good in the respective component. 
    So each component has at least one triangle that is also good in $\calT$. 
    Hence $\calT$ contains two good triangles, which completes the induction step.

@@

\begin{remark}
    Not all triangulable sets admit a triangulation that is shellable. 
    Moreover, even if a set admits a shellable triangulation, not all of its triangulations are shellable. For example, if we extend the non-shellable triangulation of the tetrahedron from~\cite[Example 8.9]{ziegler1995lectures} to a triangulation of a hypertetrahedron by suspending it from a new point $v_\star$, then the resulting new triangulation is non-shellable and coincides with the patch around $v_\star$.
    This demonstrates that patches around boundary simplices are not necessarily shellable when the space dimension $n$ is larger than three.
\end{remark}
\begin{remark}
    Not all triangulable sets admit a triangulation that is also shellable. 
    Moreover, even if a set admits a shellable triangulation, not all of its triangulations might be shellable. 
    We refer to~\cite[Example 8.9]{ziegler1995lectures} for an examples of non-shellable triangulations of cubes and tetrahedra in three dimensions. 
\end{remark}

@@

\begin{proposition}\label{proposition:starreflection}
    Let $\calT$ be a triangulation of an $n$-dimensional domain. 
    Let $S \in \calT$ be an inner simplex of dimension $k < n$,
    let $T \in \patch_{\calT}(S)$ be of dimension $n$,
    and let 
    \begin{align*}
        U := \overline{ |\patch_{\calT}(S)| \setminus T },
        \qquad 
        {\Gamma_1} := U \cap T,
        \qquad 
        {\Gamma_2} := \overline{ \partial T \setminus \partial U }.
    \end{align*}
    The following holds, where the constants on the right-hand sides are as stated in the proof. 
    \begin{enumerate}
    \item 
    There exists a bi-Lipschitz mapping
    \begin{align*}
        \Xi_{1} : T \rightarrow U
    \end{align*}
    which is the identity along ${\Gamma_1} = T \cap U$. 
    At any point, the singular values $\sigma_1 \geq \dots \geq \sigma_n$ of its Jacobian satisfy 
    \begin{gather*}
        \sigma_1 \leq \Cfive{n}{k}(\calT),
        \quad 
        \sigma_2,\dots,\sigma_n \leq \Cfiveprime{n}{k}(\calT),
        \quad 
        | \det\Jacobian\Xi_{1} |      \leq \Cdetfive{n}{k}(\calT),
        \\
        \sigma_n \leq \Csechs{n}{k}(\calT),
        \quad 
        \sigma_1^{-1},\dots,\sigma_{n-1}^{-1} \leq \Csechsprime{n}{k}(\calT),
        \quad 
        | \det\Jacobian\Xi_{1}^{-1} | \leq \Cdetsechs{n}{k}(\calT).
    \end{gather*}
%     and which satisfies the upper bounds 
%     \begin{gather*}
%         \matnorm{ \Jacobian\Xi_{1} }          \leq \Cfive{n}{k}(\calT),
%         \quad 
%         \matnorm{ \Jacobian\Xi_{1}^{-1} }     \leq \Csechs{n}{k}(\calT),
%         \\
%         | \det\Jacobian\Xi_{1} |      \leq \Cdetfive{n}{k}(\calT),
%         \quad 
%         | \det\Jacobian\Xi_{1}^{-1} | \leq \Cdetsechs{n}{k}(\calT).
%     \end{gather*}
    \item 
    There exists a bi-Lipschitz mapping
    \begin{align*}
        \Xi_{2} : \underlying{\patch_{\calT}(S)} \rightarrow U
    \end{align*}
    which is the identity along $\partial U \setminus \partial T$.
    At any point, the singular values $\sigma_1 \geq \dots \geq \sigma_n$ of its Jacobian satisfy 
    \begin{gather*}
        \sigma_1 \leq \Csieb{n}{k}(\calT),
        \quad 
        \sigma_2,\dots,\sigma_n \leq \Csiebprime{n}{k}(\calT),
        \quad 
        | \det\Jacobian\Xi_{2} |      \leq \Cdetsieb{n}{k}(\calT),
        \\
        \sigma_n \leq \Cacht{n}{k}(\calT),
        \quad 
        \sigma_1^{-1},\dots,\sigma_{n-1}^{-1} \leq \Cachtprime{n}{k}(\calT),
        \quad 
        | \det\Jacobian\Xi_{2}^{-1} | \leq \Cdetacht{n}{k}(\calT).
    \end{gather*}
%     and which satisfies the upper bounds 
%     \begin{gather*}
%         \matnorm{ \Jacobian\Xi_{2} }          \leq \Csieb{n}{k}(\calT),
%         \quad 
%         \matnorm{ \Jacobian\Xi_{2}^{-1} }     \leq \Cacht{n}{k}(\calT),
%         \\
%         | \det\Jacobian\Xi_{2} |      \leq \Cdetsieb{n}{k}(\calT),
%         \quad 
%         | \det\Jacobian\Xi_{2}^{-1} | \leq \Cdetacht{n}{k}(\calT).
%     \end{gather*}
    \end{enumerate}
\end{proposition}
\begin{proof}
    We derive the estimate in several steps. 
    In what follows, we use the notation $\hat z$ for the normalization of any vector $z \in \bbR^{n}$.
    \begin{itemize}
        \item 
        Without loss of generality, the barycenter $z_{S}$ of $S$ is the origin. 
        % Let $I := U \cap T$, which is the union of $k$ different faces of $T$, with $1 \leq k \leq n$.
        %We let ${\Gamma_2} \subseteq \partial T$ be the union of the remaining $n+1-k$ faces of $T$.
        %So ${\Gamma_2$} is within the boundary of the star $|\patch_{\calT}(S)|$ and is disjoint from $S$.
        We fix the subsimplex $S' \subseteq T$ that is complementary to $S$.
        Note that ${\Gamma_2}$ is the union of exactly those faces of $T$ that contain $S'$,
        whereas ${\Gamma_1} = U \cap T$ is the union of exactly those faces of $T$ that contain $S$.

@@

        % TODO Check that $\calS'$ is a triangulation...

@@

        \begin{align}
            \sigma_{\max} = \sqrt{ 2\tan(\beta)^2 + 1 + 2 \tan(\beta) \sqrt{ \tan(\beta)^2 + 1 } } = \sqrt{ 1 + \tan^{2}(\beta) } + \tan(\beta) \label{math:theta_max}

@@

            \sigma_{\min} = \sqrt{ 2\tan(\beta)^2 + 1 - 2 \tan(\beta) \sqrt{ \tan(\beta)^2 + 1 } } = \sqrt{ 1 + \tan^{2}(\beta) } - \tan(\beta) \label{math:theta_min}

@@

        Recall that $K \in \calR$ is obtained from $T$ via barycentric subdivisions, first at $z_{S}$ and then at $z_{S'}$. 
        Let $F_{z} \subseteq K$ be the face opposite to the vertex $z_{S'}$,
        which must be contained in some face of $T$. 
        Since $S'$ has dimension $n-k-1$, the vector $(n-k) h_{z}$ is the height vector of that face of $T$. 
        Thus 
        \begin{align*}
            \frac{ \vecnorm{ b_{z} } }{ \vecnorm{ h_{z} } }
            = 
            \frac{ \vecnorm{ b_{z} } }{ (n-k)^{-1} \| (n-k) h_{z} \| }
            = 
            \frac{ \vecnorm{ z_{S'} } }{ (n-k)^{-1} \| (n-k) h_{z} \| }
            \leq 
            (n-k) \aspectratio(T)
%             = 
%             \frac{ \vecnorm{ b_{z} } \vol(F_{z}) }{ (n-k)^{-1} n \vol(T) }
%             \leq 
%             \frac{ n-k }{n!}
%             \cdot 
%             \frac{ \diam(T)^{n} }{ \vol(T) }
            .
        \end{align*}
        This establishes bounds on the singular values of the transformation. 
        We abbreviate 
        \begin{align}\label{math:upper_bound_for_theta}
            B_{T} 
            := 
%             \sqrt{ 1 + \tan^{2}(\beta) } - \tan(\beta) 
%             \geq 
            \sqrt{ 
                1 + (n-k)^{2} \aspectratio(T)^{2} % \frac{ \diam(T)^{2n} }{ n!^{2}\vol(T)^{2} } 
            } 
            + 
            (n-k) \aspectratio(T) % \frac{ \diam(T)^{n} }{ n! \vol(T) }            
            .
        \end{align}

@@

        \begin{gather}

@@

            1 + \frac{3}{2} (k+1) \aspectratio(T) 
			\label{math:phi_max}

@@

            2 + 3 (k+1) \aspectratio(T) 
			\label{math:phi_min}
        \end{gather}

@@

            \Psi : |\calR'| \rightarrow \underlying{\calK}.

@@

        % TODO \mwl{Here we need to show that it is indeed a ball. It must be a domain triangulation.}

@@

            \frac{ \| x_K \| }{ \| x_{{G}} \| }

@@

            \frac{ \| x_{G} \| }{ \| x_{K} \| }

@@

        The vector $a$ is the difference between $\hat h_{G}$ and $\hat h_{K}$ after both are rescaled such that they lie on the affine hyperplane with normal vector $\hat x$. That difference has length at most $\tan(\alpha_{G}) + \tan(\alpha_{K})$.
        We calculate 
        \begin{align*}
            \|a\|
            = 
            \left\|
            \frac{\|x_{G}\|}{\|h_{G}\|} \hat h_{{G}} 
            - 
            \frac{\|x_{K}\|}{\|h_{K}\|} \hat h_K
            \right\|
            &
            \leq 
            \frac{\|x_{G}\|}{\|h_{G}\|} 
            + 
            \frac{\|x_{K}\|}{\|h_{K}\|} 
            % \\&
            \leq 
            (k+1) 
            \left( 1 + B_{T} \right)
            \sup_{ K \in \patch_{\calT}(S) }
            \aspectratio(K) % \frac{ \diam(K)^{n} }{ n! \vol(K) }
            .
        \end{align*}
        In summary, using that $\sqrt{s^2+t^2} \leq s + t$ for any $s,t \geq 0$, we find the upper bounds 
        \begin{align*}
            \matnorm{ \Jacobian\Psi(x) }
            &\leq  
            (k+1)B_{T}
            \diameterratio(\calT)
            \aspectratio(T) % \frac{ \diam(T)^{n} }{ n! \vol(T) }
            \cdot 
			\left( 
                1 
				+ 
				(k+1) 
				\left( 1 + B_{T} \right)
				\aspectratio(\calT) 
            \right)
			,
            \\
            \matnorm{ \Jacobian\Psi(z)^{-1} }
            &\leq  
            (k+1)
            \diameterratio(\calT)
            \aspectratio(\calT) 
            \cdot 
			\left( 
                1 
				+ 
				(k+1) 
				\left( 1 + B_{T} \right)
				\aspectratio(\calT) 
            \right)
			,
            \\
            \det\Jacobian\Psi(x)
            &\leq 
            \left( 
            (k+1)B_{T}
            \diameterratio(\calT)
            \aspectratio(T) % \frac{ \diam(T)^{n} }{ n! \vol(T) }
            \right)^{n},
            \\
            \det\Jacobian\Psi^{-1}(x)
            &\leq 
            \left( 
            (k+1)
            \diameterratio(\calT)
            \aspectratio(\calT) 
            \right)^{n}
        \end{align*}
        We have shown that $\Psi$ and $\Psi^{-1}$ are piecewise Lipschitz with respect to some essentially disjoint closed partitions over their respective domains.
        % It remains to show both are Lipschitz, which we illustrate for $\Psi$, the argument for $\Psi^{-1}$ being analogous. 
        % Rademacher's theorem then shows that both have essentially bounded weak Jacobians. 
        Every function that is piecewise Lipschitz with respect to the triangulation $\calK$ must be Lipschitz over all of $\underlying{\calK}$, since there exists a bi-Lipschitz transformation from $\underlying{\calK}$ onto the unit ball.
        % We use a technique similar to Lemma~\ref{lemma:stars_are_shellable}. 
        % After application of a piecewise linear Lipschitz mapping, 
        % we can assume that $\underlying{\calK}$ is convex without loss of generality. 
        % Any function piecewise Lipschitz with respect to $\calK$ must be Lipschitz over $\underlying{\calK}$. 
        % \mwl{TODO: not enough. Instead, use the homeomorphism onto a ball and estimate its Lipschitz constant.}

        \item 
        We define a mapping $ \Xi_{1} : T \rightarrow U = \overline{\patch_{\calT}(S) \setminus T}$ by setting 
        \begin{align*}
            \Xi_{1} := \Psi \circ \Theta.
        \end{align*}
        It preserves the interface from $T$ onto $U$. 
        Let $\sigma_1(x) \geq \dots \geq \sigma_n(x)$ be the singular values of its Jacobian at $x \in T$. 
        Combining the previous estimates for $\Psi$ and $\Theta$ with interlacing inequalities for the singular values of product matrices provides 
        \begin{align*}
            \sigma_1(x) = \matnorm{ \Jacobian \Xi_{1} }
            &\leq 
            \Cfive{n}{k}(\calT)
            :=
            (k+1) 
            \diameterratio(\calT)
            \aspectratio(\calT)
            B_{T}^{2}
            \Big( 
                1 
                + 
                (k+1) 
                \left( 1 + B_{T} \right)
                \aspectratio(\calT)
            \Big)            
            ,
            \\
            \sigma_2(x), \dots, \sigma_n(x) 
            &\leq 
            \Cfiveprime{n}{k}(\calT)
            :=
            (k+1)B_{T}
            \diameterratio(\calT)
            \aspectratio(T) 
			\cdot 
			B_T
            \\
            \sigma_n(x)^{-1} = \matnorm{ \Jacobian \Xi_{1}^{-1} }
            &\leq
            \Csechs{n}{k}(\calT)
            :=
            (k+1) 
            \diameterratio(\calT)
            \aspectratio(\calT)
            B_{T}
            \Big( 
                1 
                + 
                (k+1) 
                \left( 1 + B_{T} \right)
                \aspectratio(\calT)
            \Big)
            \\
            \sigma_1(x)^{-1}, \dots, \sigma_{n-1}(x)^{-1}
            &\leq 
            \Csechsprime{n}{k}(\calT)
            :=
            (k+1)
            \diameterratio(\calT)
            \aspectratio(\calT) 
            \cdot 
			B_T
            .
        \end{align*}
        We get the determinant estimates
        \begin{align*}
            |\det(\Xi_{1})|
            &\leq 
            \Cdetfive{n}{k}(\calT)
            :=
            (k+1)^{n}
            \diameterratio(\calT)^{n}
            B_{T}^{n}
            \aspectratio(\calT)^{n}
            ,
            \\
            |\det(\Xi_{1}^{-1})|
            &\leq 
            \Cdetsechs{n}{k}(\calT)
            :=
            (k+1)^{n}
            \diameterratio(\calT)^{n}
            \aspectratio(\calT)^{n}
            .
        \end{align*}
        This completes the desired estimates for $\Xi_{1}$.
